---
title: "Confirmatory factor analysis - Testing a tau-congeneric measurement model"
format: html
citation: 
    type: "document"
    title: |
        R Code for the Development and Validation of Measurement Instruments in the Social Sciences: Psychometric Analyses (Dimensionality, Reliability, Measurement Invariance)
    container-title: GESIS, Cologne. Data File Version 1.0.0, 
    author:
      - Melanie Viola Partsch
      - Katharina Groskurth
      - Matthias Blümke
      - Isabelle Schmidt
    doi: 10.7802/1.1982
    issued: 2020
---



Load the required packages
```{r}
#| label: setup
library(rio)
library(lavaan)

packageVersion("rio")
packageVersion("lavaan")
```

Read the example data.

```{r}
#| label: read_data
example1 <- rio::import("example1.sav")
```

Specify and estimate tau-congeneric measurement model of 3 correlated factors using robust maximum likelihood estimation. 

```{r}
#| label: model
EXT_3cF_TC <- "
  SOC =~ NA*bfi2_1 + bfi2_16R + bfi2_31R + bfi2_46
  ASS =~ NA*bfi2_6 + bfi2_21 + bfi2_36R + bfi2_51R
  ENL =~ NA*bfi2_11R + bfi2_26R + bfi2_41 + bfi2_56
  SOC ~~ 1*SOC
  ASS ~~ 1*ASS
  ENL ~~ 1*ENL"
EXT_3cF_TC.M1 <- cfa(EXT_3cF_TC, data = example1, estimator = "MLR")
```

If data is categorical specify estimator = "WLSMV" and additionally specify  `ordered = [character vector of variable names]`.

Default for missing data handling is listwise deletion. For the use of FIML additionally specify `missing = "ML"`

If an attempt should be made to mimic the Mplus program, specify `mimic = "Mplus"`

```{r}
#| label: summary
summary(EXT_3cF_TC.M1, standardized = TRUE, fit.measures = TRUE)
```
