---
title: "Confirmatory factor analysis - Testing a tau-equivalent measurement model"
format: html
author:
    - Melanie Viola Partsch
    - Katharina Groskurth
    - Matthias Blümke
    - Isabelle Schmidt
keywords:
  - Survey Data
  - Measurement
  - Validity
license: "CC BY-NC 4.0"
citation: 
    type: "document"
    title: |
        R Code for the Development and Validation of Measurement Instruments in the Social Sciences: Psychometric Analyses (Dimensionality, Reliability, Measurement Invariance)
    container-title: GESIS, Cologne. Data File Version 1.0.0, 
    author:
      - Melanie Viola Partsch
      - Katharina Groskurth
      - Matthias Blümke
      - Isabelle Schmidt
    doi: 10.7802/1.1982
    issued: 2020
---



Load the required packages
```{r}
#| label: setup
library(rio)
library(lavaan)

packageVersion("rio")
packageVersion("lavaan")
```

Read the example data.

```{r}
#| label: read_data
example1 <- rio::import("example1.sav")
```

Specify and estimate essentially tau-equivalent measurement model

```{r}
SOC_TE <- 'SOC =~ 1*bfi2_1 + 1*bfi2_16R + 1*bfi2_31R + 1*bfi2_46'
SOC_TE.M1<-cfa(SOC_TE, data = example1, estimator = "MLR") 
```
Estimate model using robust maximum likelihood estimation. If data is categorical specify `estimator = "WLSMV"` and additionally specify `ordered = [character vector of variable names]`.

Default for missing data handling is listwise deletion. For the use of FIML additionally specify `missing = "ML"`

If an attempt should be made to mimic the Mplus program, specify `mimic = "Mplus"`

```{r}
summary(SOC_TE.M1, standardized = TRUE, fit.measures = TRUE) 
```

Specify tau-congeneric measurement model for chi-square difference test.

```{r}
SOC_TC <- 'SOC =~ NA*bfi2_1 + bfi2_16R + bfi2_31R + bfi2_46
          SOC ~~ 1*SOC'
SOC_TC.M1<-cfa(SOC_TC, data = example1, estimator = "MLR") 

```
Estimate model using robust maximum likelihood estimation. If data is categorical specify `estimator = "WLSMV"` and additionally specify `ordered = [character vector of variable names]`.

Default for missing data handling is listwise deletion. For the use of FIML additionally specify `missing = "ML"`

If an attempt should be made to mimic the Mplus program, specify mimic = "Mplus"

```{r}
summary(SOC_TC.M1, standardized = TRUE, fit.measures = TRUE)
```

Conduct chi-square difference test between essentially tau-equivalent and tau-congeneric model

```{r}
chisq.diff <- fitMeasures(SOC_TE.M1)[["chisq.scaled"]] - fitMeasures(SOC_TC.M1)[["chisq.scaled"]]
df.diff <- fitMeasures(SOC_TE.M1)[["df.scaled"]] - fitMeasures(SOC_TC.M1)[["df.scaled"]]
pchisq(chisq.diff, df.diff, lower.tail = FALSE)
```
