---
title: "Measurement invariance testing"
format: html
author:
    - Melanie Viola Partsch
    - Katharina Groskurth
    - Matthias Blümke
    - Isabelle Schmidt
keywords:
  - Survey Data
  - Measurement
  - Validity
license: "CC BY-NC 4.0"
citation: 
    type: "document"
    title: |
        R Code for the Development and Validation of Measurement Instruments in the Social Sciences: Psychometric Analyses (Dimensionality, Reliability, Measurement Invariance)
    container-title: GESIS, Cologne. Data File Version 1.0.0, 
    author:
      - Melanie Viola Partsch
      - Katharina Groskurth
      - Matthias Blümke
      - Isabelle Schmidt
    doi: 10.7802/1.1982
    issued: 2020
---

Load the required packages
```{r}
#| label: setup
library(rio)
library(lavaan)

packageVersion("rio")
packageVersion("lavaan")
```

Read the example data.

```{r}
#| label: read_data
example1 <- rio::import("example1.sav")
```

Specify model for test of configural measurement invariance

```{r}
EXT_3cF <- "
  SOC =~ NA*bfi2_1 + bfi2_16R + bfi2_31R + bfi2_46
  ASS =~ NA*bfi2_6 + bfi2_21 + bfi2_36R + bfi2_51R
  ENL =~ NA*bfi2_11R + bfi2_26R + bfi2_41 + bfi2_56
  SOC ~~ 1*SOC
  ASS ~~ 1*ASS
  ENL ~~ 1*ENL"
```

estimate model for configural measurement invariance testing across females and males

```{r}
EXT_3cF.CON <- cfa(EXT_3cF, data = example1, estimator = "MLR", group = "sex")
```

Estimate model using robust maximum likelihood estimation. If data is categorical specify `estimator = "WLSMV"` and additionally specify `ordered = [character vector of variable names]`.

Default for missing data handling is listwise deletion. For the use of FIML additionally specify `missing = "ML"`

```{r}
summary(EXT_3cF.CON, standardized = TRUE, fit.measures = TRUE)
```

Specify model for test of metric and scalar measurement invariance (free estimation of variances of latent variables in all but first group)

```{r}
EXT_3cF_fLV <- "
  SOC =~ NA*bfi2_1 + bfi2_16R + bfi2_31R + bfi2_46
  ASS =~ NA*bfi2_6 + bfi2_21 + bfi2_36R + bfi2_51R
  ENL =~ NA*bfi2_11R + bfi2_26R + bfi2_41 + bfi2_56
  SOC ~~ c(1,NA)*SOC
  ASS ~~ c(1,NA)*ASS
  ENL ~~ c(1,NA)*ENL"
```

estimate model for metric measurement invariance testing across females and males.

```{r}
EXT_3cF.MET <- cfa(EXT_3cF_fLV,
    data = example1, estimator = "MLR", group = "sex",
    group.equal = c("loadings")
)
```
Estimate model using robust maximum likelihood estimation. If data is categorical specify `estimator = "WLSMV"` and additionally specify `ordered = [character vector of variable names]`.

Default for missing data handling is listwise deletion. For the use of FIML additionally specify `missing = "ML"`

```{r}
summary(EXT_3cF.MET, standardized = TRUE, fit.measures = TRUE)
```

estimate model for scalar measurement invariance testing across females and males

```{r}
EXT_3cF.SCA <- cfa(EXT_3cF_fLV, data = example1, estimator = "MLR", group = "sex", group.equal = c("loadings", "intercepts"))
```

Estimate model using robust maximum likelihood estimation. 
If data is categorical specify `estimator = "WLSMV"` and additionally specify `ordered = [character vector of variable names]`.
Default for missing data handling is listwise deletion. For the use of FIML additionally specify `missing = "ML"`

```{r}
summary(EXT_3cF.SCA, standardized = TRUE, fit.measures = TRUE)
```


extract CFI, RMSEA and SRMR for the configural, metric and scalar measurement model

```{r}
EXT_3cF_fitCon<-fitMeasures(EXT_3cF.CON, c("cfi.scaled", "rmsea.scaled", "srmr"))
EXT_3cF_fitMet<-fitMeasures(EXT_3cF.MET, c("cfi.scaled", "rmsea.scaled", "srmr"))
EXT_3cF_fitSca<-fitMeasures(EXT_3cF.SCA, c("cfi.scaled", "rmsea.scaled", "srmr"))
```

calculate the delta/difference in CFI, RMSEA and SRMR for configural vs. metric and metric vs. scalar.

```{r}
EXT_3cF_deltaCFI_CtoM <- round((EXT_3cF_fitCon[1] - EXT_3cF_fitMet[1]), digits = 3)
EXT_3cF_deltaCFI_MtoS <- round((EXT_3cF_fitMet[1] - EXT_3cF_fitSca[1]), digits = 3)
EXT_3cF_deltaRMSEA_CtoM <- round((EXT_3cF_fitCon[2] - EXT_3cF_fitMet[2]), digits = 3)
EXT_3cF_deltaRMSEA_MtoS <- round((EXT_3cF_fitMet[2] - EXT_3cF_fitSca[2]), digits = 3)
EXT_3cF_deltaSRMR_CtoM <- round((EXT_3cF_fitCon[3] - EXT_3cF_fitMet[3]), digits = 3)
EXT_3cF_deltaSRMR_MtoS <- round((EXT_3cF_fitMet[3] - EXT_3cF_fitSca[3]), digits = 3)

```

## Display results

CFI, RMSEA and SRMR for the configural, metric and scalar measurement model

```{r}
EXT_3cF_fitCon
EXT_3cF_fitMet
EXT_3cF_fitSca
```

delta/differences between fit indices from configural to metric model

```{r}
EXT_3cF_deltaCFI_CtoM
EXT_3cF_deltaRMSEA_CtoM
EXT_3cF_deltaSRMR_CtoM
```

delta/differences between fit indices from metric to scalar model

```{r}
EXT_3cF_deltaCFI_MtoS
EXT_3cF_deltaRMSEA_MtoS
EXT_3cF_deltaSRMR_MtoS
```
