---
title: "Principal component analysis"
format: html
citation: 
    type: "document"
    title: |
        R Code for the Development and Validation of Measurement Instruments in the Social Sciences: Psychometric Analyses (Dimensionality, Reliability, Measurement Invariance)
    container-title: GESIS, Cologne. Data File Version 1.0.0, 
    author:
      - Melanie Viola Partsch
      - Katharina Groskurth
      - Matthias Blümke
      - Isabelle Schmidt
    doi: 10.7802/1.1982
    issued: 2020
---



Load the required packages
```{r}
#| label: setup
library(rio)
library(psych)

packageVersion("rio")
packageVersion("psych")
```

Read the example data.

```{r}
#| label: read_data
example1 <- rio::import("example1.sav")
```

Build character vector with variable names for use below

```{r}
ext <- c(
    "bfi2_1", "bfi2_6", "bfi2_11R", "bfi2_16R", "bfi2_21", "bfi2_26R",
    "bfi2_31R", "bfi2_36R", "bfi2_41", "bfi2_46", "bfi2_51R",
    "bfi2_56"
)
```

Conduct parallel analysis, print eigen values and display screeplot

Function with arguments specified below corresponds to default 
20 resampled or simulated random datasets are used and the empirical eigen values are compared to the mean of the resampled or simulated eigenvalues; Suitable for the use of continuous data

Alternatives to default settings:  
Increase number of resampled or simulated datasets (e.g., to n.iter = 100) and choose a quantile of resampled/simulated data to compare the empirical eigenvalues against (e.g., `quant = .95`; the suggested number of components likely might decrease with a high value for quant)

When using categorical data: specify `cor = "poly"`.

Info: Plots can be saved as image or pdf via click on Export

```{r}
fa.parallel(example1[ext],
    fa = "pc",
    n.iter = 20, # quant = .95,
    cor = "cor"
)$fa.values
```


## Conduct PCA

Function below specifies the following:  

- extraction of 3 components

- no rotation (i.e., initial (orthogonal) component solution)

- suitable for the use of continuous data


Alternative specification of arguments:  

- orthogonal rotation: specify rotation = "varimax"

- oblique rotation: specify `rotation = "oblimin"` or `rotation = "promax"` (the latter includes Kaiser normalization)

- obtain oblique component scores based on pattern matrix: `oblique.scores = FALSE`

- when using categorical data: specify `cor = "poly"`

```{r}
myPCA <- principal(example1[ext],
    nfactors = 3,
    rotate = "none",
    oblique.scores = T,
    cor = "cor"
)
myPCA
```
